---
title: "week 5 code port_folio"
author: "Abhi"
date: "June 16, 2019"
output: html_document
---

## Matrices and Dataframes

# to create a dataframe/matrix

# cbind(a, b, c)	# this combines vectors as columns in a matrix
```{r}
cbind(3:8, 9:14, 15:20)
```

# rbind(a, b, c)	 # Combine vectors as rows in a matrix	
```{r}
rbind(3:8, 9:14, 15:20)
```


# matrix(x, nrow, ncol, byrow)    #	Creates a matrix from a vector x




# data.frame()	 # Creates a dataframe from named columns	

```{r}
data.frame("age" = c(45, 55), 
sex = c("m", "f"))
```
```{r}
setwd("C:/Users/A/Desktop/Late Spring @HU/ANLY 506-51- B- Exploratory Data Analytics-Olga/Text books")
chicago <- readRDS("chicago.rds")
```
# matrix and dataframe functions

# head() function # lists the first few observations of the dataframe ,default no.= 6
``` {r}
head(chicago)
```
# tail() function #  lists the bottom few observations of the dataframe ,default no.= 6
``` {r}
tail(chicago)
```
# View() functions # displays the full object in a new tab
``` {r}
View(chicago)
```
# nrow() funtion  # counts the total no. of rows 
``` {r}
nrow(chicago)
```
# ncol() funtion  # counts the total no. of columns
``` {r}
ncol(chicago)
```
# dim() function # the dimension functions give the total number of rows and columns
``` {r}
dim(chicago)
```
# str() function # it shows the structure of the data set ie. the no. of rows and columns and few observations from each column and also the data type for that column
``` {r}
str(chicago)
```
# summary() function # this gives the 5-number statistical summary for each of the variables and also specifies the no. of NA's in each of them separately

``` {r}
summary(chicago)
```
# names() function   # gives the column names for the dataframe
``` {r}
names(chicago)
```
# mean() function  # gives the arithmetic mean for the chosen object
``` {r}
mean(chicago$tmpd)
```
# Adding a new column to a dataframe
 
dataframe$new <- c(a,b,c....)  # adds a new column called 'new' to the data frame with the data point as a,b,c....

``` {r}
chicago$new <- c(1:6940)
```
# Changing column names

names(dataframe)[3] <- "z"  # this changes the name of the 3rd column to 'z' 

names(dataframe)[names(dataframe) == "old.name"] <- "new.name" #  better approach to avoid errors
``` {r}
names(chicago)[names(chicago) == "new"] <- "very_new"
str(chicago)
```

# Slicing the dataframe

1.dataframe[3,] # returns the observations in 3rd row and all the columns

2.dataframe[ , 5] # returns the observations for all the rows but only the 5th column

3.dataframe[40:100,3:8] # returns the observations in rows 40th to 100 for columns 3rd to 8th
``` {r}
chicago[20:25,2:4]
```
4.dataframe_new <- dataframe[dataframe$a < 50 & dataframe$b > 40]  # using logical operator with specific conditions on columns "a" and "b" by the use of AND logic
``` {r}
chicago_new <- chicago["tmpd" > 50 & "dptp" <40]
head(chicago_new)
```

5.subset() function for slicing

subset(x = dataframe,
      subset = a< 20 &
               b == 5 &
               c >= 10)   # slicing out a subset from the main dataframe using logical conditions on differen columns/variable wiz. a,b,c

``` {r}
a = subset( chicago,subset = chicago$tmpd >50 & chicago$dptp < 40)
head(a)
```               
               
               