---
title: "week 3 code_portfolio"
author: "Abhi"
date: "June 16, 2019"
output: html_document
---
#Workspace management functions

# ls() function  # this displays all the objects in current workspace
```{r}
ls()
```
rm(x,y,z,.....) # this removes the objects named 'x','y','z'... from the current workspace
rm(list = ls()) # removes all the objects from the workspace

getwd() function # displays the current working directory
``` {r}
getwd()
```

1.setwd(file = "dir) # changes the working directory to a desired location

2.list.files() # returns the list of all files in the current working directory

3.save(a, b, .., file = "myimage.RData)	# this saves objects a, b, . to myimage.RData

4.save.image(file = "myimage.RData")	#Saves all objects in your workspace to myimage.RData

5.load(file = "myimage.RData")	# this Loads objects in the file myimage.RData




##loading a text file into R



read.table(file = "mydata.txt", sep, header)	    # this reads a ".txt"" file named "mydata.txt", and we can set how columns are separated with "sep" ie. sep = "," for comma-delimited files, and sep = "\t" for tab-delimited files and header = TRUE sets a specific header column






# Exporting objects as text file




 write.table(x = object,
            file = "mytext.txt",  # Saving the file as mytext.txt
            sep = "\t" /",")   # saving as a comma-separated file or a tab-delimited file




    
# Loading data from a website

# fromweb <- read.table(file = 'the url link of the website',sep = '\t', header = TRUE)
                      
                      
            