---
title: "week 8"
author: "Abhi"
date: "July 03, 2019"
output: html_document
---
```{r warning=FALSE , message=FALSE , error=FALSE}
library(tidyverse)
```
# Visualizing Distributions

## For Categorical variables

We can use a simple barplot for the categorical variable "cut"
```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut))
```
Here the height of the bars show the count for each variable

## For continous variables

```{r}
ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = carat), binwidth = 0.5)
```
Here a histogram is used to represent the continous variable "carat" and the width of the intervals is set by the arguement "binwith"

### Using geom_freqpoly()

```{r}
smaller <- diamonds %>% 
filter(carat < 3)
ggplot(data = smaller, mapping = aes(x = carat, colour = cut)) +
geom_freqpoly(binwidth = 0.1)
```

Here the overlapping lines are easier to understand



# The Lattice system
By the lattice system plots are created by a single call and much information about multi-dimensional data can be represented

```{r warning=FALSE , message=FALSE , error=FALSE}
library(lattice)
```

```{r}
state <- data.frame(state.x77, region = state.region)
xyplot(Life.Exp ~ Income | region, data = state, layout = c(4, 1))
```

This plot contains 4 panels- one for each geographical reason showing the relation between income and life expectancy



