---
title: "week 7_code_portfolio"
author: "Abhi"
date: "July 2, 2019"
output: html_document
---

```{r warning=FALSE , message=FALSE , error=FALSE}
library(tidyverse)
```

# Graphing template 

###  ggplot(data = <DATA>) +  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>)) # <DATA> represents the data to be plotted , <GEOM_FUNCTION> refers to the Geometric Function used for plotting and aes refers to the aesthetics like size,shape and colour of the points

``` {r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

### Here we use data = mpg which is build in R, on the x-axis "displ" which is the engine size in litres and "hwy" which is the fuel efficiency on the highway


## Adding another variable 

###By colour

We can add one more variable from the data set in the form of different colors for different values


```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

Here we add the variable "class" represented by the different colors.
ggplot will add a unique color to each unique value of the class variable and will also automatically display a legend for the same.


###By size

Here we add the variable "class" represented by the different sizes of the plotting points.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = class))
```

### By transparency of the points (alpha)

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class))
```

### By shape

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```

### By facets

We can add extra discrete variables by adding facets or subplots that each display one subset of the data

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

## Geometric Objects

Geometric objects or Geom are used by the plot to represent data

### Geom as a point

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

### Geom as a smooth line

```{r}
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

### Adding variables by line type

```{r}
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))
```

### Adding variables by colour of lines

```{r}
ggplot(data = mpg) +
  geom_smooth(
    mapping = aes(x = displ, y = hwy, color = drv),
    show.legend = FALSE)
    ```
    
### Adding multiple Geoms to the plot

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
```


## Colour and Fill 
We can use colour and fill aesthethics in a bar plot to simultaneously show 2 variables

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity))
```


## Coordinate systems in ggplot

### coord_flip() 
This command just switches the x and y axes

```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot()
```
A normal boxplot
 
 
```{r}
 ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot() +
  coord_flip()
```
A horizontal boxplot by coord_flip()

  
  
  
